On the specific recursive structure of iterated parametricity

### 2025-09-28 paper/paper.tex
paper: typesetting improvements, dce

\newcommandx{\cohtwo}[9][2,3,4,5,6,7,8,9]{
  \ensuremath{\mathsf{\color{chestnut}{coh2}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4][#5][#6][#7]}^{\joinargs[#8]}}
  \setkeys*[bonak]{coh#1}{#9}
}

@@

% Drop the universe letter in fullframe for all remaining tables
\renewcommandx{\fullframe}[3][1,2,3]{\prim{fullframe}[][#2][][][#3]}
\def\lab{tab:frames}
\input{tab-frames.tex}

@@

% Coherence conditions for frame, layer, and painting
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohp}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}

@@

Defining $\painting[n][p]$, in addition to the type dependency in $\framep[n][p]$, requires the definition of $\painting[n-1][p-1]$, thus also $\framep[n-1][p-1]$ by typing dependency, as well as $\restrf[n][p-1]$. Since $\painting[n][p]$ also depends on $\painting[n][p+1]$ when $p<n$, these dependencies are actually between $\painting[n][p,...,n]$, $\painting[n-1][p,...,n-1]$, $\framep[n][p,...,n]$, $\framep[n-1][p,...,n-1]$ and $\restrf[n][p,...,n]$. Also, for the case $p=n$, this also requires the assumption of an inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally, $\framep[n][p,...,n]$, $\framep[n-1][p,...,n-1]$ and $\restrf[n][p,...,n-1]$ respectively depend on the earlier values $\framep[n][0,...,p-1]$, $\framep[n-1][0,...,p-1]$ and $\restrf[n][0,...,p-1]$.  Below, we use dashed underline and dashed arrows for dependencies in subconstructions, here in the construction of $\framep$ at level $n-1$ and $\restrftype$ at level $n$:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    & & \painting[n][p,...,n] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
    \udensdash{\framep}^{n,[0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
    \udensdash{\framep}^{n,[p,...,n]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
    \painting[n-1][0,...,p-1] \arrow[d, dotted] & &
    \painting[n-1][p,...,n-1] \arrow[ll, dotted] \arrow[d, dotted] \\
    \framep[n-1][0,...,p-1] &
    \udensdash{\restrf}^{n,[0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
    \framep[n-1][p,...,n-1] \arrow[ll, dotted, bend left=15] &
    \udensdash{\restrf}^{n,[p,...,n-1]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
  \end{tikzcd}
\end{center}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][p,...,n]         &                                              & :      & Type_{l+1}                                                                                   \\
    \frametype[n][]                &                                              & \defeq & unit                                                                                         \\
    \frametype[n][p,...,n]         &                                              & \defeq & Type_l \times \frametype[n][p+1,...,n-1]                                                     \\
    \\
    \paintingtype[n][p,...,n]      & (\framep[n][p,...,n]:\frametype[n][p,...,n]) & :      & Type_{l+1}                                                                                   \\
    \paintingtype[n][]             & \star                                        & \defeq & unit                                                                                         \\
    \paintingtype[n][p,...,n](p<n) & (\framep[n][p,...,n])                        & \defeq & (\framep[n][p] \rightarrow Type_l) \times \paintingtype[n][p+1,...,n](\framep[n][p+1,...,n]) \\
    \\
    \restrftype[n][p,...,n]        &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]:\frametype[n-1][0,...,p-1]                               \\
                \painting[n-1][0,...,p-1]:\paintingtype[n-1][0,...,p-1](\framep[n-1][0,...,p-1]) \\
                \framep[n-1][p,...,n-1]:\frametype[n-1][p,...,n-1]                               \\
                \painting[n-1][p,...,n-1]:\paintingtype[n-1][p,...,n-1](\framep[n-1][p,...,n-1]) \\
                \restrf[n][0,...,p-1]:\restrftype[n][0,...,p-1]
                \left(\begin{array}{l}
                    \framep[n-1][0,...,p-1]   \\
                    \painting[n-1][0,...,p-1] \\
                  \end{array}\right)                                                  \\
              \end{array}\right) : Type_{l}
    }                                                                                                                                                                                     \\
    \restrftype[n][]               &
    \left(\begin{array}{l}
              \framep[n-1][0,...,n-1]   \\
              \painting[n-1][0,...,n-1] \\
              \star                     \\
              \star                     \\
              \restrp[n][0,...,n-1]     \\
            \end{array}\right)   & \defeq                                       & unit                                                                                                      \\
    \restrftype[n][p,...,n-1]      &
    \left(\begin{array}{l}
              \framep[n-1][0,...,p-1]   \\
              \painting[n-1][0,...,p-1] \\
              \framep[n-1][p,...,n-1]   \\
              \painting[n-1][p,...,n-1] \\
              \restrf[n][0,...,p-1]     \\
            \end{array}\right)
                                   & \defeq                                       &
    \begin{array}{l}
      \framep[n][p]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]   \\
                \painting[n-1][0,...,p-1] \\
                \restrf[n][0,...,p-1]     \\
              \end{array}\right)
      \rightarrow \framep[n-1][p] \\
      \times                      \\
      \restrftype[n][p+1,...,n]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p]       \\
                \painting[n-1][0,...,p]     \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
                \restrf[n][0,...,p]         \\
              \end{array}\right)
    \end{array}
  \end{array}
\end{equation*}

>>>>

@@

\def\lab{tab:frames}
\input{tab-frames.tex}

@@

% Coherence conditions for frame, layer, and painting
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohp}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}

@@

Defining $\painting[n][p]$, in addition to the type dependency in $\framep[n][p]$, requires the definition of $\painting[n-1][p-1]$, thus also $\framep[n-1][p-1]$ by typing dependency, as well as $\restrf[n][p-1]$. Since $\painting[n][p]$ also depends on $\painting[n][p+1]$ when $p<n$, these dependencies are actually between $\painting[n][p,...,n]$, $\painting[n-1][p,...,n-1]$, $\framep[n][p,...,n]$, $\framep[n-1][p,...,n-1]$ and $\restrf[n][p,...,n]$. Also, for the case $p=n$, this also requires the assumption of an inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally, $\framep[n][p,...,n]$, $\framep[n-1][p,...,n-1]$ and $\restrf[n][p,...,n-1]$ respectively depend on the earlier values $\framep[n][0,...,p-1]$, $\framep[n-1][0,...,p-1]$ and $\restrf[n][0,...,p-1]$.  Below, we use dashed underline and dashed arrows for dependencies in subconstructions, here in the construction of $\framep$ at level $n-1$ and $\restrftype$ at level $n$:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    & & \painting[n][p,...,n] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
    \udensdash{\framep[n][0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
    \udensdash{\framep[n][p,...,n]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
    \painting[n-1][0,...,p-1] \arrow[d, dotted] & &
    \painting[n-1][p,...,n-1] \arrow[ll, dotted] \arrow[d, dotted] \\
    \framep[n-1][0,...,p-1] &
    \udensdash{\restrf[n][0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
    \framep[n-1][p,...,n-1] \arrow[ll, dotted, bend left=15] &
    \udensdash{\restrf[n][p,...,n-1]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
  \end{tikzcd}
\end{center}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][p,...,n]         &                                              & :      & Type_{l+1}                                                                                   \\
    \frametype[n][]                &                                              & \defeq & unit                                                                                         \\
    \frametype[n][p,...,n]         &                                              & \defeq & Type_l \times \frametype[n][p+1,...,n-1]                                                     \\
    \\
    \paintingtype[n][p,...,n]      & (\framep[n][p,...,n]:\frametype[n][p,...,n]) & :      & Type_{l+1}                                                                                   \\
    \paintingtype[n][]             & \star                                        & \defeq & unit                                                                                         \\
    \paintingtype[n][p,...,n](p<n) & (\framep[n][p,...,n])                        & \defeq & (\framep[n][p] \rightarrow Type_l) \times \paintingtype[n][p+1,...,n](\framep[n][p+1,...,n]) \\
    \\
    \restrftype[n][p,...,n]        &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]:\frametype[n-1][0,...,p-1]                               \\
                \painting[n-1][0,...,p-1]:\paintingtype[n-1][0,...,p-1](\framep[n-1][0,...,p-1]) \\
                \framep[n-1][p,...,n-1]:\frametype[n-1][p,...,n-1]                               \\
                \painting[n-1][p,...,n-1]:\paintingtype[n-1][p,...,n-1](\framep[n-1][p,...,n-1]) \\
                \restrf[n][0,...,p-1]:\restrftype[n][0,...,p-1]
                \left(\begin{array}{l}
                    \framep[n-1][0,...,p-1]   \\
                    \painting[n-1][0,...,p-1] \\
                  \end{array}\right)                                                  \\
              \end{array}\right) : Type_{l}
    }                                                                                                                                                                                     \\
    \restrftype[n][]               &
    \left(\begin{array}{l}
              \framep[n-1][0,...,n-1]   \\
              \painting[n-1][0,...,n-1] \\
              \star                     \\
              \star                     \\
              \restrp[n][0,...,n-1]     \\
            \end{array}\right)   & \defeq                                       & unit                                                                                                      \\
    \restrftype[n][p,...,n-1]      &
    \left(\begin{array}{l}
              \framep[n-1][0,...,p-1]   \\
              \painting[n-1][0,...,p-1] \\
              \framep[n-1][p,...,n-1]   \\
              \painting[n-1][p,...,n-1] \\
              \restrf[n][0,...,p-1]     \\
            \end{array}\right)
                                   & \defeq                                       &
    \begin{array}{c}
      \framep[n][p]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]   \\
                \painting[n-1][0,...,p-1] \\
                \restrf[n][0,...,p-1]     \\
              \end{array}\right)
      \rightarrow \framep[n-1][p] \\
      \times                      \\
      \restrftype[n][p+1,...,n]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p]       \\
                \painting[n-1][0,...,p]     \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
                \restrf[n][0,...,p]         \\
              \end{array}\right)
    \end{array}
  \end{array}
\end{equation*}
